---
title: "Final Report"
output: 
  html_document:
    toc: true
    toc_float: true
---

<br/>  

#### Motivation: Provide an overview of the project goals and motivation.

Running a marathon requires an immense amount of time, effort, and dedication. First-time marathon runners are forced to make decisions with little knowledge of what to expect. Being ill-prepared and over-training are two common experiences that occur during training and on race day. Emma, one of our group members, ran the New York City Marathon (NYCM) this past fall, which was her first marathon. After completing the marathon on November 6th, Emma was prepared to start training for the next one. Therefore, we are interested in understanding the impact of external factors, such as temperature, wind, elevation, etc., on Emma's speed and performance during the marathon, while also considering how training changed over time in order to improve her training for next marathon.

<br/>  

#### Related Work: Anything that inspired you, such as a paper, a web site, or something we discussed in class.

This year, the NYCM was considered the hottest marathon since 1986. With temperatures in the mid-70s for most of the race, participants were encouraged to focus on hydrating and maintaining electrolytes in order to maintain their speed and finish the race. Most runners - beginners and experts alike - were struggling with the heat during race day, causing adverse health outcomes and decreased performance ([Runner's World Article](https://www.runnersworld.com/news/a41956910/strava-data-2022-nyc-marathon/)). Emma's experiences before and during the race inspired us to investigate her marathon training data. During the marathon, Emma felt the effects of the heat on her performance, and watched numerous people collapse mid-race. Emma also had her own battles with training prior to the marathon most likely due to over-training.

Upon receiving Emma's data from Strava.com, we noticed that the data was clean and organized, allowing for easy generation of visuals. After reviewing the data, we recognized that we could conduct analyses from different vantage points, such as reviewing the average elements for each run or using per-second data to generate interactive maps. Inspired by our in-class work using `ggplot`, `plot.ly`, and additional visualization techniques, we are able to comprehensively investigate Emma's training and race day data. Lastly, we hope to use modeling to understand the interactions of different factors on Emma's performance, which will hopefully inform her training efforts in the future. 

<br/>

#### Initial Questions: What questions are you trying to answer? How did these questions evolve over the course of the project? What new questions did you consider in the course of your analysis?



<br/>  

#### Data: Source, scraping method, cleaning, etc.

Data was collected using the Garmin Forerunner 245 activity watch from July 23, 2020 to November 6, 2022. Collected data was uploaded to Strava.com after each activity. After granting permission, personal data was downloaded from Strava.com. Summary data for each activity post was saved in the form of a `.csv` file (imported using `tidyverse`) and per-second data for each activity was compiled and saved in a `.fit` file (imported using `FITfileR`). 

##### *Importing Data*

First, we imported the summary data of all of Emma's training runs from July 23, 2020 to November 6, 2022. In order to clean this data, we:

1. Sorted data to include only information from the time that Emma started training for the NYCM, which was on April 4, 2022
2. Selected the necessary variables

```{r, load packages}
library(tidyverse)
library(FITfileR)
library(dplyr)
library(patchwork)
library(leaflet)
```


```{r}
training_summary <- read_csv("activities/activities.csv") %>% 
  janitor::clean_names() %>%
  filter(activity_type == "Run") %>%
  select(activity_id, activity_date, activity_name, activity_description, elapsed_time_6, distance_7, max_heart_rate_8, relative_effort_9, max_speed, average_speed, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, average_cadence, average_heart_rate, calories, weather_temperature, dewpoint, humidity, wind_speed) %>%
  filter(activity_id >= 6910869137) %>%
  separate(activity_date, c("month_date", "year", "time"), sep = ", ") %>%
  mutate(
    date = str_c(month_date, year, sep = " "),
    date = as.Date(date, format = "%b%d%Y")
  ) %>%
  select(-month_date, -year) %>%
  mutate(
    elapsed_time_min = elapsed_time_6 / 60,
    distance_km = distance_7,
    max_heart_rate = max_heart_rate_8,
    relative_effort = relative_effort_9
  ) %>%
  select(-elapsed_time_6, -distance_7, -max_heart_rate_8, -relative_effort_9)
activity_summaries =
  training_summary %>%
  select(activity_id, date, time, activity_name, activity_description)
tidy_training = 
  training_summary %>%
  select(-activity_name, -activity_description) %>%
  select(activity_id, date, time, distance_km, elapsed_time_min, max_speed, average_speed, max_heart_rate, average_heart_rate, relative_effort, everything()) %>% 
  filter(date != "2022-03-31", 
         date != "2022-04-03")
```

Since each Garmin activity was downloaded as an individual FIT file, a function was created to iterate across the multiple activities and compile a single data frame. As shown below, the function was created to read each activity FIT file into R and pull out the records (which included data on latitude and longitude, distance, speed, heart rate, and cadence), and then each second of activity was merged into a single data frame. This function was applied to a list of all file names. A `for` loop was used to generate a list of data frames, and then rows were merged using `bind_rows` to create the final data frame, `marathon_df`. 

##### *Data Cleaning*
After importing, the data frame was then cleaned to include activities from the start of marathon training (March 31, 2022) to the day of the New York City Marathon (November 6, 2022). Within the 'activities' data set, there were 335 observations and 84 variables. The data was filtered to only include observations from 'activity_type' that were classified as 'Run'. From this, 22 variables were selected to be included in this data set: activity_id, activity_date, activity_name, activity_description, elapsed_time_6, distance_7, max_heart_rate_8, relative_effort_9, max_speed, average_speed, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, average_cadence, average_heart_rate, calories, weather_temperature, dewpoint, humidity,and wind_speed. 

Due to Emma's marathon training beginning March 31st, 2022, the 'activity_id' was used in order to filter any training prior to this. Following this, the variable 'activity_date' was separated into two variables:the date on which the run occurred ('date') as well as the time when the run began ('time'). The 'elapsed_time_6' variable was converted from seconds to minutes ('elapsed_time_min'). In addition to this, the variables names for distance_7, max_heart_rate_8, and relative_effort_9 were changed to distance_km, max_heart_rate, and relative_effort.

A new data frame, 'activities_summaries' was created in order to include the written accounts from each of Emma's runs, including variables 'activity_name' and 'activity_description'.

The final cleaned data set, 'tidy_training', included 110 observations and 21 variables. Variables included were:activity_id, date, time, distance_km, elapsed_time_min, max_speed, average_speed, max_heart_rate, average_heart_rate, relative_effort, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, average_cadence, calories, weather_temperature, dewpoint, humidity, and wind_speed. 

#### Exploratory Analysis: Visualizations, summaries, and exploratory statistical analyses. Justify the steps you took, and show any major changes to your ideas.

* Mileage over time (Intro) - Individual monthly or weekly mileage
* Rounded mileage to closest mile, density plot of how many times you ran a certain mileage or violin 
* Temperature vs average speed/average heart rate/relative effort (all of these)
 -- * Elevation vs average speed/relative effort 
* Mapping running trails (plotly/gis(?)) - only marathon route, map cadence/heart rate
* Predict mph in marathon using temperature, humidity, relative effort, etc. 

Side note: Rest days are not in strava data - create data set with all dates. we make a variable that is yes/no for rest days.

#### Mileage over time

```{r}
tidy_training = tidy_training %>% 
  mutate(week = as.numeric(strftime(date, format = "%V")) - 13) %>% 
  group_by(week) %>%
  mutate(weekly_distance_km = sum(distance_km))

tidy_training %>%
  ggplot(aes(x = week, y = weekly_distance_km)) +
  geom_point(color = c("#FFA500")) +
  geom_line(color = c("#FFA500")) + 
  labs(
    title = "Weekly Distance Ran by Emma (km)",
    x = "Week", 
    y = "Distance (km)") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5))
```

#### Mileage Density Plot

```{r}
tidy_training %>% 
  ggplot(aes(x = distance_km), color = c("#FFA500")) +
  geom_density(alpha = 0.5, fill = c("#FFA500")) + 
  labs(
    title = "Density of Daily Distances Ran by Emma (km)",
    x = "Distance (km)", 
    y = "Density") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5))
```


#### Choose Your Graph (Drop down list of Y and X)
Temperature vs average speed/average heart rate/relative effort (all of these)

##### Temperature vs. Average Speed

```{r}
tidy_training %>% 
  ggplot(aes(x = weather_temperature, y = average_speed)) + 
  geom_point(color = c("#FFA500")) +
  geom_smooth(se = FALSE, color = "dark grey") + 
    labs(
    title = "Temperature (C) vs. Emma's Average Speed",
    x = "Temperature (C)", 
    y = "Average Speed") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5))
```

##### Temperature as a Predictor for Emma's Performance

```{r}
temp_avg_speed = 
  tidy_training %>% 
  ggplot(aes(x = weather_temperature, y = average_speed)) + 
  geom_point(color = c("#FFA500")) +
  geom_smooth(se = FALSE, color = "dark grey") + 
    labs(
    x = "Temperature (C)", 
    y = "Average Speed") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Temperature (C) vs. Emma's Average Speed")

temp_avg_hrt = 
  tidy_training %>% 
  ggplot(aes(x = weather_temperature, y = average_heart_rate)) + 
  geom_point(color = c("#FFA500")) +
  geom_smooth(se = FALSE, color = "dark grey") + 
    labs(
    x = "Temperature (C)", 
    y = "Average Heart Rate (BPM)") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Temperature (C) vs.\n Emma's Average Heart Rate (BPM)")

temp_rel_effort = 
  tidy_training %>% 
  ggplot(aes(x = weather_temperature, y = relative_effort)) + 
  geom_point(color = c("#FFA500")) +
  geom_smooth(se = FALSE, color = "dark grey") + 
    labs(
    x = "Temperature (C)", 
    y = "Relative Effort") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Temperature (C) vs.\n Emma's Relative Effort")

temp_avg_speed /
(temp_avg_hrt | temp_rel_effort)
```

##### Humidity as a Predictor for Emma's Performance

```{r}
humid_avg_speed = 
  tidy_training %>% 
  ggplot(aes(x = humidity, y = average_speed)) + 
  geom_point(color = c("#FFA500")) +
  geom_smooth(se = FALSE, color = "dark grey") + 
    labs(
    x = "Humidity (%)", 
    y = "Average Speed") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Humidity (%) vs. Emma's Average Speed")

humid_avg_hrt = 
  tidy_training %>% 
  ggplot(aes(x = humidity, y = average_heart_rate)) + 
  geom_point(color = c("#FFA500")) +
  geom_smooth(se = FALSE, color = "dark grey") + 
    labs(
    x = "Humidity (%)", 
    y = "Average Heart Rate (BPM)") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Humidity (%) vs.\n Emma's Average Heart Rate (BPM)")

humid_rel_effort = 
  tidy_training %>% 
  ggplot(aes(x = humidity, y = relative_effort)) + 
  geom_point(color = c("#FFA500")) +
  geom_smooth(se = FALSE, color = "dark grey") + 
    labs(
    x = "Humidity (%)", 
    y = "Relative Effort") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + 
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Humidity (%) vs.\n Emma's Relative Effort")

humid_avg_speed /
(humid_avg_hrt | humid_rel_effort)
```

*Map of Marathon Route with Averages*
NOTE: Install Leaflet Prior to Running

Loading and Cleaning the Data
```{r}
marathon_day <- read_csv("activities/marathon_day.csv")

marathon_day =
marathon_day %>% 
mutate(
  distance_m = distance*0.000621371
) %>% 
mutate(group = case_when(
    between(distance_m, 0,0.9999) ~ "Start",
    between(distance_m, 1, 1.9999) ~ "1",
    between(distance_m, 2, 2.9999) ~ "2",
    between(distance_m, 3, 3.9999) ~ "3",
    between(distance_m, 4, 4.9999) ~ "4",
    between(distance_m, 5, 5.9999) ~ "5",
    between(distance_m, 6, 6.9999) ~ "6",
    between(distance_m, 7, 7.9999) ~ "7",
    between(distance_m, 8, 8.9999) ~ "8",
    between(distance_m, 9, 9.9999) ~ "9",
    between(distance_m, 10, 10.9999) ~ "10",
    between(distance_m, 11, 11.9999) ~ "11",
    between(distance_m, 12, 12.9999) ~ "12",
    between(distance_m, 13, 13.9999) ~ "13",
    between(distance_m, 14, 14.9999) ~ "14",
    between(distance_m, 15, 15.9999) ~ "15",
    between(distance_m, 16, 2.9999) ~ "16",
    between(distance_m, 17, 17.9999) ~ "17",
    between(distance_m, 18, 18.9999) ~ "18",
    between(distance_m, 19, 19.9999) ~ "19",
    between(distance_m, 20, 20.9999) ~ "20",
    between(distance_m, 21, 21.9999) ~ "21",
    between(distance_m, 22, 22.9999) ~ "22",
    between(distance_m, 23, 23.9999) ~ "23",
    between(distance_m, 24, 24.9999) ~ "24",
    between(distance_m, 25, 25.9999) ~ "25",
    between(distance_m, 26, 26.9999) ~ "26-Finish")) 
```

Recode

```{r}
marathon_speed =
marathon_geo %>% 
  group_by(group) %>% 
  mutate(
    speed_km = 
  enhanced_speed*2.237)
```

Aggregate Data
```{r}
agg = aggregate(marathon_speed,
                by = list(marathon_geo$group),
                FUN = mean)
```

Aggregate Leaflet Plot

```{r}
leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(data = agg,
             lng = ~longitude,
             lat = ~latitude,
             label = ~Group.1,
             radius = 7,
             color = "orange",
             stroke = TRUE, fillOpacity = 0.75,
             popup = ~paste("Mile:", Group.1,
                            "<br>Speed:", speed_km,
                            "<br>HR:", heart_rate,
                             "<br>Cadence:", cadence,
                            "<br>Altitude:", enhanced_altitude)) %>% 
  addProviderTiles(providers$CartoDB.Positron)
```

