---
title: "Regression Models: Relative Effort"
output: 
  html_document:
    toc: true
    code_folding: hide
    toc_float: true
---

<br/>  

### Regression Models: Relative Effort

```{r, message = FALSE}
library(tidyverse)
library(FITfileR)
library(dplyr)
library(patchwork)
library(leaflet)
library(modelr)

training_raw <- read_csv("activities/activities.csv") %>% 
  janitor::clean_names()

training_summary = 
  training_raw %>% 
  filter(activity_type == "Run") %>%
  select(activity_id, activity_date, activity_name, activity_description, elapsed_time_6, distance_7, max_heart_rate_8, relative_effort_9, max_speed, average_speed, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, average_cadence, average_heart_rate, calories, weather_temperature, dewpoint, humidity, wind_speed) %>%
  filter(activity_id >= 6910869137) %>%
  separate(activity_date, c("month_date", "year", "time"), sep = ", ") %>%
  mutate(
    date = str_c(month_date, year, sep = " "),
    date = as.Date(date, format = "%b%d%Y")
  ) %>%
  select(-month_date, -year) %>%
  mutate(
    elapsed_time_min = elapsed_time_6 / 60,
    distance_km = distance_7,
    max_heart_rate = max_heart_rate_8,
    relative_effort = relative_effort_9
  ) %>%
  select(-elapsed_time_6, -distance_7, -max_heart_rate_8, -relative_effort_9)
activity_summaries =
  training_summary %>%
  select(activity_id, date, time, activity_name, activity_description)

tidy_training = 
  training_summary %>%
  select(-activity_name, -activity_description) %>%
  select(activity_id, date, time, distance_km, elapsed_time_min, max_speed, average_speed, max_heart_rate, average_heart_rate, relative_effort, everything()) %>% 
  filter(date != "2022-03-31", 
         date != "2022-04-03")
```

Additionally, we hypothesize that Emma's relative effort is dependent on temperature. To test this hypothesis, we built the following models: 

* Relative effort vs temperature 
* Relative effort vs dewpoint
* Relative effort vs humidity
* Relative effort vs elevation gain and wind speed, and all interactions
* Relative effort vs temperature and dewpoint, and all interactions

```{r, warning = FALSE, message = FALSE, error = FALSE}
fit1 = lm(relative_effort ~ weather_temperature, data = tidy_training)
fit2 = lm(relative_effort ~ dewpoint, data = tidy_training)
fit3 = lm(relative_effort ~ humidity, data = tidy_training)
fit4 = lm(relative_effort ~ (elevation_gain + wind_speed)^3, data = tidy_training)
fit5 = lm(relative_effort ~ (weather_temperature + dewpoint + humidity)^4, data = tidy_training)
```


### Cross-Validation: Relative Effort

We followed the same cross validation process as above. 

```{r, warning = FALSE, message = FALSE, error = FALSE}
cv_df = 
  crossv_mc(tidy_training, 100)

cv_df = cv_df %>% 
  mutate(
    mod1 = map(.x = train, ~lm(relative_effort ~ weather_temperature, data = .x)), 
    mod2 = map(.x = train, ~lm(relative_effort ~ dewpoint, data = .x)), 
    mod3 = map(.x = train, ~lm(relative_effort ~ humidity, data = .x)), 
    mod4 = map(.x = train, ~lm(relative_effort ~ (elevation_gain + wind_speed)^3, data = .x)),
    mod5 = map(.x = train, ~lm(relative_effort ~ (weather_temperature + dewpoint + humidity)^4, data = .x))) %>% 
  mutate(
    rmse_mod1 = map2_dbl(.x = mod1, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_mod2 = map2_dbl(.x = mod2, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_mod3 = map2_dbl(.x = mod3, .y = test, ~rmse(model = .x, data = .y)),
    rmse_mod4 = map2_dbl(.x = mod4, .y = test, ~rmse(model = .x, data = .y)),
    rmse_mod5 = map2_dbl(.x = mod5, .y = test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(), 
    names_to = "model", 
    values_to = "rmse", 
    names_prefix = "rmse_") %>% 
  ggplot(aes(x = model, y = rmse, color = c("#FFA500"))) +
  geom_boxplot(color = c("#FFA500")) + 
  labs(
    title = "Root Mean Squared Errors Distributions for Emma's Relative Effort", 
    x = "Fitted Model", 
    y = "RMSE")  + 
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(labels = c("mod1" = "Temperature Model", 
                              "mod2" = "Dewpoint Model", 
                              "mod3" = "Humidity Model",
                              "mod4" = "Elevation, Wind, \n Interactions",
                              "mod5" = "Temperature, Dewpoint, \n Humidity, Interactions"))

summary(fit4)
```

Based on the distributions depicted above, we can conclude that the model with elevation, wind, and interactions is able to generate better predictions than the other models with environmental factors. The p-value for this model is < 2.2 e-16. As it is below the significance level, we can conclude that this model is a good predictor for average speed. 
