---
title: "Regression Models: Speed"
output: 
  html_document:
    toc: true
    code_folding: hide
    toc_float: true
---

<br/>  

### Regression Models: Speed

```{r, message = FALSE}
library(tidyverse)
library(FITfileR)
library(dplyr)
library(patchwork)
library(leaflet)
library(modelr)

training_raw <- read_csv("activities/activities.csv") %>% 
  janitor::clean_names()

training_summary = 
  training_raw %>% 
  filter(activity_type == "Run") %>%
  select(activity_id, activity_date, activity_name, activity_description, elapsed_time_6, distance_7, max_heart_rate_8, relative_effort_9, max_speed, average_speed, elevation_gain, elevation_loss, max_grade, average_grade, max_cadence, average_cadence, average_heart_rate, calories, weather_temperature, dewpoint, humidity, wind_speed) %>%
  filter(activity_id >= 6910869137) %>%
  separate(activity_date, c("month_date", "year", "time"), sep = ", ") %>%
  mutate(
    date = str_c(month_date, year, sep = " "),
    date = as.Date(date, format = "%b%d%Y")
  ) %>%
  select(-month_date, -year) %>%
  mutate(
    elapsed_time_min = elapsed_time_6 / 60,
    distance_km = distance_7,
    max_heart_rate = max_heart_rate_8,
    relative_effort = relative_effort_9
  ) %>%
  select(-elapsed_time_6, -distance_7, -max_heart_rate_8, -relative_effort_9)
activity_summaries =
  training_summary %>%
  select(activity_id, date, time, activity_name, activity_description)

tidy_training = 
  training_summary %>%
  select(-activity_name, -activity_description) %>%
  select(activity_id, date, time, distance_km, elapsed_time_min, max_speed, average_speed, max_heart_rate, average_heart_rate, relative_effort, everything()) %>% 
  filter(date != "2022-03-31", 
         date != "2022-04-03")
```

Based on Emma's recount of personal experience, she perceives environmental factors to be significant predictors of her performance. More specifically, weather conditions such as temperature, humidity, dewpoint, and windspeed affect Emma's perceived running experience. Additionally, due to the irregular environmental conditions on race day, we generated several models to investigate which environmental factors would be best predictors for Emma's performance. We hypothesize that Emma's average speed is dependent on temperature. To test this hypothesis, we built the following models: 

* Average speed vs temperature 
* Average speed vs dewpoint
* Average speed vs humidity
* Average speed vs elevation gain and wind speed, and all interactions
* Average speed vs temperature and dewpoint, and all interactions

```{r, warning = FALSE, message = FALSE, error = FALSE}
fit1 = lm(average_speed ~ weather_temperature, data = tidy_training)
fit2 = lm(average_speed ~ dewpoint, data = tidy_training)
fit3 = lm(average_speed ~ humidity, data = tidy_training)
fit4 = lm(average_speed ~ (elevation_gain + wind_speed)^3, data = tidy_training)
fit5 = lm(average_speed ~ (weather_temperature + dewpoint)^3, data = tidy_training)
```

### Cross-Validation: Speed

We can now perform a comparison between the five models in terms of cross-validated prediction error. To do this, we first create a cross-validation data frame `cv_df`. Contained within `cv_df` is 2 side-by-side list columns with the testing and training data split pairs, with a corresponding ID number for each pair. Since we are able to apply resample objects directly into `lm` functions, we can skip the extra step of converting the testing and training data into tibbles. Next, we apply the `map` function to map each regression model to the training data. Finally, we can compute the root mean squared errors (RMSEs) for each model by applying the `map2_dbl` function to the testing data. 

```{r, warning = FALSE, message = FALSE, error = FALSE}
cv_df = 
  crossv_mc(tidy_training, 100)

cv_df = cv_df %>% 
  mutate(
    mod1 = map(.x = train, ~lm(average_speed ~ weather_temperature, data = .x)), 
    mod2 = map(.x = train, ~lm(average_speed ~ dewpoint, data = .x)), 
    mod3 = map(.x = train, ~lm(average_speed ~ humidity, data = .x)), 
    mod4 = map(.x = train, ~lm(average_speed ~ (elevation_gain + wind_speed)^3, data = .x)),
    mod5 = map(.x = train, ~lm(average_speed ~ (weather_temperature + dewpoint)^3, data = .x))) %>% 
  mutate(
    rmse_mod1 = map2_dbl(.x = mod1, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_mod2 = map2_dbl(.x = mod2, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_mod3 = map2_dbl(.x = mod3, .y = test, ~rmse(model = .x, data = .y)),
    rmse_mod4 = map2_dbl(.x = mod4, .y = test, ~rmse(model = .x, data = .y)),
    rmse_mod5 = map2_dbl(.x = mod5, .y = test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(), 
    names_to = "model", 
    values_to = "rmse", 
    names_prefix = "rmse_") %>% 
  ggplot(aes(x = model, y = rmse, color = c("#FFA500"))) +
  geom_boxplot(color = c("#FFA500")) + 
  labs(
    title = "Root Mean Squared Errors Distributions for Emma's Average Speed", 
    x = "Fitted Model", 
    y = "RMSE")  + 
  theme(axis.line = element_line(color = "grey"), 
        panel.background = element_blank(), 
        legend.position = "none", 
        panel.grid.major = element_line(color = "light grey", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(labels = c("mod1" = "Temperature Model", 
                              "mod2" = "Dewpoint Model", 
                              "mod3" = "Humidity Model",
                              "mod4" = "Elevation, Wind, \n Interactions",
                              "mod5" = "Temperature, Dewpoint, \n Interactions"))
  
summary(fit1) %>% 
  broom::tidy() %>% 
  filter(term == "weather_temperature") %>% 
  mutate(
    lower_ci = exp(estimate - 1.96*std.error),
    upper_ci = exp(estimate + 1.96*std.error)
  ) %>% 
  select(term, estimate, lower_ci, upper_ci, p.value) %>% 
  knitr::kable(digits = 3, 
               col.names = c("Term", "Estimate", "Lower CI", "Upper CI", "P-value"))
```

Based on the distributions depicted above, we can conclude that the model with temperature is able to generate better predictions than the other models with environmental factors. The p-value for this model is 0.03101. As it is below the significance level of 0.05, we can conclude that this model is a good predictor for average speed. 
